Planetary perturbers: Flaring star-planet interactions in Kepler and TESS

### 2023-10-23 src/tex/ms.tex
[showyourwork] overleaf sync

% Abstract with filler text
\begin{abstract}
    In many star-planet systems discovered so far, the innermost planet orbits within only a few stellar radii. In these systems, planets \textcolor{red}{could} become in-situ probes of the extended stellar magnetic field. Because they disturb the field as they move, they are expected to trigger \textcolor{red}{flares in the corona. Potential differences to the energies and morphologies of intrinsic flares are poorly constrained.} However, as we expect planet-induced flares to \textcolor{red}{correlate} with the planet's orbital period, we can identify \textcolor{red}{them} from a clustering of flares in phase with the planet's orbit. We used the excellent phase coverage from Kepler and the Transiting Exoplanet Survey Satellite to find flaring star-planet systems, compile a catalog of all their flares, and measure how much they cluster in orbital phase. In the 1811 searched systems, we found 25 single stars with three or more flares each. We quantified the significance of the clustering in each system, and compared it against the theoretically expected power of magnetic interaction that leads to planet-induced flaring. Most systems do not show any clustering, consistent with low expected power. Those we expect to show clustering fall on two branches. An inactive one, without any signs of clustering, and a tentative active one, where the clustering becomes more pronounced as the expected power of interaction increases. \textcolor{red}{The flares in HIP 67522 are prominently clustered ($p<0.006$). This young Hot Jupiter system is the most promising candidate for magnetic star-planet interaction in our sample.}

@@

\subsection{Magnetic star-planet interactions}
\label{sec:intro:mspi}
Magnetic SPI can occur if the planet is orbiting within the Alfv\'en radius of the star at least part of the time~\citep{preusse2006magnetic, cohen2011dynamics}. The Alfv\'en radius is the radius at which the stellar wind velocity, which increases with increasing distance from the star, exceeds the Alfv\'en velocity of the magnetized plasma. Beyond this radius, this plasma is disconnected from the star. Alfv\'en waves set off by \textcolor{red}{a planet that crosses the magnetic field} cannot propagate back to the star if the wind carries the plasma away faster than these waves can travel. In the sub-Alfv\'enic regime, the wave can reach the star, and deposit its energy in its atmosphere. \citet{lanza2012starplanet, lanza2018closeby}, and \citet{zarka2007plasma} and \citet{saur2013magnetic} suggest different mechanisms for when and how the energy transport and dissipation takes place, and how much energy can be transferred and dissipated at all. They all lack observational constraints from large samples for calibration, which we aim to provide in this work.

\subsection{Searching for global changes in activity}
\label{sec:intro:global}
\textcolor{red}{Even if the instrument's sensitivity is sufficiently high, measuring magnetic star planet-interaction remains notoriously difficult. In individual systems,} a non-detection may not mean an absence of the interaction, but merely a temporary cessation~\citep{shkolnik2005hot, shkolnik2008nature}. Statistical studies searching for changes in global chromospheric and coronal activity indicators in star-planet systems with close-in planets relative to planet-less stars are inconclusive~\citep{kashyap2008extrasolar,scharf2010possible, shkolnik2013ultraviolet, france2018farultraviolet, viswanath2020statistical, krejcova2012evidence, miller2015comprehensive, poppenhaeger2010coronal}. In these studies, it is challenging both to define a control sample to measure the activity against, and to quantify how the way we detect planets introduces selection bias in the activity measure. It is, for instance, easier to detect a planet around a magnetically inactive star, because both radial velocity and transit detections will be less affected by stellar variability~\citep{poppenhaeger2011correlation}, but that does not imply that SPI quenches activity. Further, tidal interactions can induce \textcolor{red}{a} global increase in activity~\citep{ilic2022tidal}, which adds a physical ambiguity to the interpretation~(see also Section~\ref{sec:discussion:tidal}).

@@

The absolute power of flaring SPI is difficult to constrain. We do not know how energetic planet-induced flares can be. The estimates for dissipated energy~\citep{lanza2018closeby} are in the ballpark of the regime of superflares\textcolor{red}{, with energies released at optical wavelengths $\geq10^{33}\,$erg~\citep{schaefer2000superflares}, that can be} observed by Kepler and TESS. However, on the one hand, the planet might be triggering flares prematurely, which would be of lower energy~\citep{loyd2023flares}, and end up below our detection threshold. On the other hand, the energy need not be dissipated continuously, particularly in the stretch-and-break mechanism proposed by~\citep{lanza2012starplanet}, so an individual flare's energy might be higher. We also do not know whether flares triggered by interactions with a companion would \textcolor{red}{unfold faster or slower than} intrinsic flares. We will assume that, since they would be caused by magnetic reconnection triggered by the planet in the stellar corona, they should still have the same physics as intrinsic flares\textcolor{red}{. They should then approximalety follow the fast-rise-exponential-decay morphology~(see Fig.~\ref{fig:flares_and_fps} for an example). Individual intrinsic flares can deviate from this template, yet still be classified as flares because flares remain the best explanation for a sudden rise and drop in stellar brightness in most main sequence dwarf stars. Many of these deviations are unexplained today~\citep{howard2022no}, so that identifying SPI flares based on morphology is not possible without additional constraints. However, if we can statistically identify enough SPI flares, as we attempt in this work, we can infer the differences between intrinsic and SPI events from the two ensembles.}

\textcolor{red}{In other words, s}ince planet-induced flares might not look any different from intrinsic flares, and we do not know how their energies will be distributed, we require a method to distinguish them without referring to their individual properties. Instead, we can look at the statistical distribution of flares in time:

@@

\subsection{Searching for local changes in activity}
\label{sec:intro:local}
In contrast to global SPI, local SPI effects are easier to tell apart from intrinsic stellar behavior because they are tied to the planet's orbital period. \textcolor{red}{Searches for an orbital modulation of polarized radio emission have produced promising, but tentative results. Close-in star-planet systems are expected to interact similarly to the Jupiter-Io system~\citep{bigg1964influence}, albeit scaled up to the high field strengths of stars~\citep{turnpenney2018exoplanetinduced}. In some detections, the planet has not been found (yet)~\citep{vedantham2020coherent}, or the data do not cover multiple orbital periods to see the signal repeat~\citep{perez-torres2021monitoring, pineda2023coherent}, or both \citep{callingham2021population}.} Phase-correlated activity, flaring and otherwise, has been searched for in individual systems. After eight years of monitoring of HD 189733, an active Hot Jupiter host, higher X-ray energy flares seem only marginally clustered in orbital phase than lower energy ones~\citep{pillitteri2022xray}. \citet{maggio2015coordinated} observed an eruption of X-ray emission when the Jupiter-sized companion was in periastron on the highly eccentric orbit around its host HD 17156. However, a similar study using chromospheric indicators in the eccentric system HD 80606 found no variation between periastron and apoastron~\citep{figueira2016activity}. In the above cases, non-detection can at least partly be attributed to poor orbital phase coverage, which also limits the significance of any positive detection. Realizing this, \citet{fischer2019timevariable} used K2 data with excellent phase coverage of the orbits of the inner \mbox{TRAPPIST-1} planets to look for clustered flaring in phase, but found none. \citet{ilin2022searching} and \citet{klein2022one} found hints of excess flaring, and chromospheric Helium emission in the young AU Mic system, respectively. This mixed picture of individual object studies calls for a more comprehensive and systematic approach.

@@

\section{Data}
\label{sec:data}
In this study, we aimed at compiling the largest possible sample to search for flaring SPI. We took the entire Planetary Systems Composite Parameters Table (PSCP\footnote{ \url{https://exoplanetarchive.ipac.caltech.edu/cgi-bin/TblView/nph-tblView?app=ExoTbls&config=PSCompPars}}, Section~\ref{sec:data:sps}) from the NASA Exoplanet Archive, as of July 2022, and scanned the systems for available Kepler and TESS light curves~(Section~\ref{sec:data:photometry}). We searched a total of 1811 systems and over 7200 light curves from both missions for flares. To calculate the clustering of \textcolor{red}{flares} in orbital phase, we further required the orbital periods of each system~(Section~\ref{sec:data:orbitalperiod}). We then estimated the theoretically expected power of interaction for each system with at least three flares. For this, we required further system parameters -- distance between planet and star, stellar rotation period, planetary radius, and stellar luminosity~(Sections~\ref{sec:data:a}-\ref{sec:data:lum}), which we mostly took from the literature. We began by compiling the table of system parameters with the PSCP, but verified, supplemented and updated the table with more up-to-date measurements. The relevant properties of the final sample of 25 systems are shown in Table~\ref{tab:maintable_lit}.

\begin{table*}
\begin{rotatetable*}
\script{paper_main_table.py}
% \rotate
 \caption{Flaring single star-planet system parameters. Obs. time is the total observing time with Kepler and TESS.}
\movetableright=-40mm

@@

The Transiting Exoplanet Survey Satellite~\citep[TESS,][]{ricker2015transiting} is an all-sky mission that began operations in 2018, \textcolor{red}{and has} completed two sky scans in summer 2022. It is observing at the time of writing, collecting nearly continuous photometric time series in the $600-1000\,$nm band for $\sim 27\,$d in each observing Sector. About $200\,000$ stars have been observed in 2-min cadence in the first two years of operations alone, with about $20,000$ targets per Sector. Out of these, from Sector 27 on, $1,000$ targets were observed at even higher 20-s cadence in each Sector. 
% 20000 per sector does not add up, because of stars observed in multiple sectors, the 20000 and 1000 figures can be found here: https://tess.mit.edu/observations/target-lists/
% extended TESS mission https://heasarc.gsfc.nasa.gov/docs/tess/extended.html

@@

% \textcolor{red}{Mid-to-late M dwarfs follow this bimodal distribution }

@@

The methods for light curve de-trending and flare finding in Kepler and TESS light curves, as well as the Anderson-Darling test, are the same as detailed in~\citet{ilin2022searching}. We briefly recap the techniques in Sections~\ref{sec:methods:flaresearch} and \ref{sec:methods:adtest}. The expected power of SPI depends on the relative velocity between the planet and the magnetic field strength in its orbit, the derivation of which we explain in 
Sections~\ref{sec:methods:relvel} and \ref{sec:methods:bfield}, respectively. We can then combine them with \textcolor{red}{the} stellar radius, planetary radius, and semi-major axis to estimate the power of SPI. We use the scaling laws for two different magnetic SPI models, which we introduce in Section~\ref{sec:methods:pspi}.

\subsection{Light curve de-trending and flare search}
\label{sec:methods:flaresearch}
To remove trends and rotational variability from the light curve without losing the flare signal, we use an empirically derived multistep process, implemented as the \texttt{custom\_detrending} method in \texttt{AltaiPony}, a flare science Python package for light curve analysis~\citep{ilin2021altaipony}. First, we apply a spline fit with a coarse sampling of 30h-averaged values to capture slow rotation with periods above multiple days and non-periodic trends. Then, we iteratively fit a series of sines to capture rotational signal on time scales down to about half a day, close to the fastest rotational signals measured in low mass stars. Eventually, we apply two Savitzky-Golay~\citep{savitzky1964smoothing} filters in sequence, with window sizes of 6h and 3h each. At this stage, we mask all data points above a $2.5 \sigma$ \textcolor{red}{threshold} (or $1.5 \sigma$ for stars like AU Mic or Proxima Cen, which are very active or have very low noise levels, or both) as flare candidates to prevent the filter from ironing out the flares. As a final step, we fit exponential functions to the edges of the light curves, if there are data points that deviate more than one standard deviation from the median value, while keeping the flare candidates masked. Transits are usually too shallow to affect either light curve de-trending or flare finding. The light curve portions around deep transits were inspected manually.

@@

We vet all flare candidates by eye, and exclude instrumental and physical false positives, such as Solar System Objects passing by the field of view~(see Fig.~\ref{fig:flares_and_fps} in the Appendix for an example). For each confirmed flare candidate, we calculate the relative amplitude $a$ and equivalent duration $ED$, defined as the flare flux $F_{\rm flare}$ over the duration of the flare, divided by the median flux $F_0$ of the star, integrated over the flare duration~\citep{gershberg1972results}:
\begin{equation}
\label{eq:ED}
ED=\displaystyle \int \mathrm dt\, \frac{F_{\rm flare}(t)}{F_0}.
\end{equation}
\textcolor{red}{In other words, equivalent duration is }the time during which the non-flaring star releases as much energy as the flare.

\subsection{Custom Anderson-Darling test}
\label{sec:methods:adtest}
For each star-planet system with three or more flares in its light curves, we test for deviations from a random distribution of flares with orbital phase. We use the same customized Anderson-Darling test as in \cite{ilin2022searching}, which is more sensitive to deviations at both ends of the distribution than the widely used Kolmogorov-Smirnov test~\citep{feigelson2012modern}. In \textcolor{red}{short}, we first take the number of flares observed in the TESS and Kepler data to calculate a base flare rate for each light curve. We then calculate how often each orbital phase has been observed, also for each light curve separately, to ensure that the different detection thresholds for TESS and Kepler do not bias the base rate. With the phase coverage and base flare rate combined, we can tell how many flares we would expect to see in any given phase bin if flares were randomly distributed. We then aggregate this number of expected flares per \textcolor{red}{orbital phase} bin into one expected distribution \textcolor{red}{(blue curves in Figs.~\ref{fig:cumdist_transiting} and \ref{fig:cumdist_rv})} by summing over all available light curves. As a last step, we compare the expected distribution to the observed one using an Anderson-Darling test. \textcolor{red}{For this, we sample 10000 times from the expected distribution the same number of flares as we observe in each system. For instance, for HIP 67522, we sample 12 flares distributed according to its blue curve in Fig.~\ref{fig:cumdist_transiting}. For each of the 10000 samples, we compute the Anderson-Darling statistic $A^2$, which yields a distribution of $A^2$ values. In a last step, we compute $A^2$ for the observed distribution, and compute the $p$-value.} We repeat this test with different phase offsets to account for remaining biases in sensitivity of the test at different phases.

@@

\subsection{Stellar magnetic fields}
\label{sec:methods:bfield}
We derive the average magnetic field strength $B$ from \textcolor{red}{Rossby number} $R$o using the empirical relation derived in \citet{reiners2022magnetism},  Table 2, in the unsaturated and saturated regimes, respectively:

@@

\begin{figure}[ht!]
    \script{paper_lx_plots.py}
    \begin{centering}
        \includegraphics[width=\linewidth]{figures/PAPER_lxlbol.png}
        \caption{
           X-ray luminosity over bolometric luminosity compared to Rossby number (\textbf{upper panel}), and average surface magnetic field strength (\textbf{lower panel}). \textcolor{red}{Blue lines denote the error bars, black lines point to the IDs of the corresponding stars.} Both relations follow those of stars not selected for hosting close-in planets, except for GJ 3082, which is underluminous for its Rossby number, and hence also its estimated magnetic field appears too strong for its $L_X/L_{\rm bol}$.
        }
        \label{fig:lx}
    \end{centering}
\end{figure}

@@

First, we estimate the power of SPI generated by the stretch-and-break mechanism using the scaling relations in~\cite{lanza2012starplanet}. \textcolor{red}{We adopt their Eqn.~45 for non-linear and axisymmetric magnetic fields, because the non-linear formulation does not assume a constant force-free parameter, and because only axisymmetric fields can produce modulation of flaring activity in phase with the planetary orbit. We note that the axisymmetry of stellar magnetic fields varies across stars~\citep[e.g.,][]{donati2009magnetic, donati2008largescale, morin2008largescale, morin2010largescale}, and also with the stellar activity cycle of individual stars~\citep[e.g.,][]{borosaikia2016solarlike, lehmann2021identifying}. In Eqn. 45, the power of SPI, $P_{\rm spi,sb}$, is}

\begin{equation}
\textcolor{red}{      P_{\rm spi,sb}=\underbrace{\frac{27\pi \lambda^2(n+1)}{16 \mu_0 n (\lambda^2 + n^2)^{1/3}}}_{\hat{C}} B_{\rm p}^{2/3} R_{\rm p}^2 B_*^{4/3}  v_{\rm rel} \left(\frac{R_*}{a}\right)^{\frac{n+11}{ 3}},}
\end{equation}

\textcolor{red}{where $\mu_0$ is the magnetic permeability of the vacuum, and $n$ and $\lambda$ are dimensionless constants, all of which we subsume under $\hat{C}$, together with other constant factors}. $B_{\rm p}$ and $B_*$ are the planetary and stellar field strengths at their poles, respectively; $v_{\rm rel}$ is the relative velocity between the stellar rotation and the planet's orbit at the semi-major axis $a$; and $R_*$ and $R_{\rm p}$ are the stellar and planetary radii, respectively. \textcolor{red}{We choose $n=0.25$ ($\lambda^2 = 1.01203$), which best reproduces observations of SPI candidate systems HD~179949, $\tau$ Boo, and HD~189733 in~\citet{lanza2012starplanet} (except for $\upsilon$ And). Since the planetary fields are unknown, we adopt $B_{\rm p}=1\,$G for simplicity.} We also do not know the stellar field strength at the pole for the vast majority of the stars, so we adopt \textcolor{red}{an average value of 15 per cent of the} surface field strength~(see Sections~\ref{sec:methods:bfield} and~\ref{sec:discussion:as}) as a proxy.

@@

\begin{equation}
\textcolor{red}{      P_{\rm spi,sb0} =\hat{C} R_{\rm p}^2 B_*^2 v_{\rm rel}  \left(\frac{R_*}{a}\right)^{\frac{n+11}{ 3}}}
\end{equation}

Third, we estimate the power of SPI generated by the Alfv\'en wing mechanism ~\citep{saur2013magnetic,kavanagh2022radio}, \textcolor{red}{ combining Eqn.~8 for the power of interaction and Eqn.~11 for the magnetopause radius of the planet from~\citet{kavanagh2022radio} to arrive at:}

\begin{equation}
\textcolor{red}{    P_{\rm spi,aw} = \frac{\pi^{1/2}}{2^{2/3}} B_{\rm p}^{2/3} R_{p}^2  R_*^2  B_*^{1/3}  v_{\rm rel}^2 a^{-2} \rho_*^{1/2} \sin^2 \theta}
\end{equation}

>>>>

% Abstract with filler text
\begin{abstract}
    In many star-planet systems discovered so far, the innermost planet orbits within only a few stellar radii. In these systems, planets could become in-situ probes of the extended stellar magnetic field. Because they disturb the field as they move, they are expected to trigger flares in the corona. Potential differences to the energies and morphologies of intrinsic flares are poorly constrained. However, as we expect planet-induced flares to correlate with the planet's orbital period, we can identify them from a clustering of flares in phase with the planet's orbit. We used the excellent phase coverage from Kepler and the Transiting Exoplanet Survey Satellite to find flaring star-planet systems, compile a catalog of all their flares, and measure how much they cluster in orbital phase. In the 1811 searched systems, we found 25 single stars with three or more flares each. We quantified the significance of the clustering in each system, and compared it against the theoretically expected power of magnetic interaction that leads to planet-induced flaring. Most systems do not show any clustering, consistent with low expected power. Those we expect to show clustering fall on two branches. An inactive one, without any signs of clustering, and a tentative active one, where the clustering becomes more pronounced as the expected power of interaction increases. The flares in HIP 67522 are prominently clustered ($p<0.006$). This young Hot Jupiter system is the most promising candidate for magnetic star-planet interaction in our sample.

@@

\subsection{Magnetic star-planet interactions}
\label{sec:intro:mspi}
Magnetic SPI can occur if the planet is orbiting within the Alfv\'en radius of the star at least part of the time~\citep{preusse2006magnetic, cohen2011dynamics}. The Alfv\'en radius is the radius at which the stellar wind velocity, which increases with increasing distance from the star, exceeds the Alfv\'en velocity of the magnetized plasma. Beyond this radius, this plasma is disconnected from the star. Alfv\'en waves set off by a planet that crosses the magnetic field cannot propagate back to the star if the wind carries the plasma away faster than these waves can travel. In the sub-Alfv\'enic regime, the wave can reach the star, and deposit its energy in its atmosphere. \citet{lanza2012starplanet, lanza2018closeby}, and \citet{zarka2007plasma} and \citet{saur2013magnetic} suggest different mechanisms for when and how the energy transport and dissipation takes place, and how much energy can be transferred and dissipated at all. They all lack observational constraints from large samples for calibration, which we aim to provide in this work.

\subsection{Searching for global changes in activity}
\label{sec:intro:global}
Even if the instrument's sensitivity is sufficiently high, measuring magnetic star planet-interaction remains notoriously difficult. In individual systems, a non-detection may not mean an absence of the interaction, but merely a temporary cessation~\citep{shkolnik2005hot, shkolnik2008nature}. Statistical studies searching for changes in global chromospheric and coronal activity indicators in star-planet systems with close-in planets relative to planet-less stars are inconclusive~\citep{kashyap2008extrasolar,scharf2010possible, shkolnik2013ultraviolet, france2018farultraviolet, viswanath2020statistical, krejcova2012evidence, miller2015comprehensive, poppenhaeger2010coronal}. In these studies, it is challenging both to define a control sample to measure the activity against, and to quantify how the way we detect planets introduces selection bias in the activity measure. It is, for instance, easier to detect a planet around a magnetically inactive star, because both radial velocity and transit detections will be less affected by stellar variability~\citep{poppenhaeger2011correlation}, but that does not imply that SPI quenches activity. Further, tidal interactions can induce a global increase in activity~\citep{ilic2022tidal}, which adds a physical ambiguity to the interpretation~(see also Section~\ref{sec:discussion:tidal}).

@@

The absolute power of flaring SPI is difficult to constrain. We do not know how energetic planet-induced flares can be. The estimates for dissipated energy~\citep{lanza2018closeby} are in the ballpark of the regime of superflares, with energies released at optical wavelengths $\geq10^{33}\,$erg~\citep{schaefer2000superflares}, that can be observed by Kepler and TESS. However, on the one hand, the planet might be triggering flares prematurely, which would be of lower energy~\citep{loyd2023flares}, and end up below our detection threshold. On the other hand, the energy need not be dissipated continuously, particularly in the stretch-and-break mechanism proposed by~\citep{lanza2012starplanet}, so an individual flare's energy might be higher. We also do not know whether flares triggered by interactions with a companion would unfold faster or slower than intrinsic flares. We will assume that, since they would be caused by magnetic reconnection triggered by the planet in the stellar corona, they should still have the same physics as intrinsic flares. They should then approximalety follow the fast-rise-exponential-decay morphology~(see Fig.~\ref{fig:flares_and_fps} for an example). Individual intrinsic flares can deviate from this template, yet still be classified as flares because flares remain the best explanation for a sudden rise and drop in stellar brightness in most main sequence dwarf stars. Many of these deviations are unexplained today~\citep{howard2022no}, so that identifying SPI flares based on morphology is not possible without additional constraints. However, if we can statistically identify enough SPI flares, as we attempt in this work, we can infer the differences between intrinsic and SPI events from the two ensembles.

In other words, since planet-induced flares might not look any different from intrinsic flares, and we do not know how their energies will be distributed, we require a method to distinguish them without referring to their individual properties. Instead, we can look at the statistical distribution of flares in time:

@@

\subsection{Searching for local changes in activity}
\label{sec:intro:local}
In contrast to global SPI, local SPI effects are easier to tell apart from intrinsic stellar behavior because they are tied to the planet's orbital period. Searches for an orbital modulation of polarized radio emission have produced promising, but tentative results. Close-in star-planet systems are expected to interact similarly to the Jupiter-Io system~\citep{bigg1964influence}, albeit scaled up to the high field strengths of stars~\citep{turnpenney2018exoplanetinduced}. In some detections, the planet has not been found (yet)~\citep{vedantham2020coherent}, or the data do not cover multiple orbital periods to see the signal repeat~\citep{perez-torres2021monitoring, pineda2023coherent}, or both \citep{callingham2021population}. Phase-correlated activity, flaring and otherwise, has been searched for in individual systems. After eight years of monitoring of HD 189733, an active Hot Jupiter host, higher X-ray energy flares seem only marginally clustered in orbital phase than lower energy ones~\citep{pillitteri2022xray}. \citet{maggio2015coordinated} observed an eruption of X-ray emission when the Jupiter-sized companion was in periastron on the highly eccentric orbit around its host HD 17156. However, a similar study using chromospheric indicators in the eccentric system HD 80606 found no variation between periastron and apoastron~\citep{figueira2016activity}. In the above cases, non-detection can at least partly be attributed to poor orbital phase coverage, which also limits the significance of any positive detection. Realizing this, \citet{fischer2019timevariable} used K2 data with excellent phase coverage of the orbits of the inner \mbox{TRAPPIST-1} planets to look for clustered flaring in phase, but found none. \citet{ilin2022searching} and \citet{klein2022one} found hints of excess flaring, and chromospheric Helium emission in the young AU Mic system, respectively. This mixed picture of individual object studies calls for a more comprehensive and systematic approach.

@@

\section{Data}
\label{sec:data}
In this study, we aimed at compiling the largest possible sample to search for flaring SPI. We took the entire Planetary Systems Composite Parameters Table (PSCP\footnote{ \url{https://exoplanetarchive.ipac.caltech.edu/cgi-bin/TblView/nph-tblView?app=ExoTbls&config=PSCompPars}}, Section~\ref{sec:data:sps}) from the NASA Exoplanet Archive, as of July 2022, and scanned the systems for available Kepler and TESS light curves~(Section~\ref{sec:data:photometry}). We searched a total of 1811 systems and over 7200 light curves from both missions for flares. To calculate the clustering of flares in orbital phase, we further required the orbital periods of each system~(Section~\ref{sec:data:orbitalperiod}). We then estimated the theoretically expected power of interaction for each system with at least three flares. For this, we required further system parameters -- distance between planet and star, stellar rotation period, planetary radius, and stellar luminosity~(Sections~\ref{sec:data:a}-\ref{sec:data:lum}), which we mostly took from the literature. We began by compiling the table of system parameters with the PSCP, but verified, supplemented and updated the table with more up-to-date measurements. The relevant properties of the final sample of 25 systems are shown in Table~\ref{tab:maintable_lit}.

\begin{table*}
\begin{rotatetable*}
\script{paper_main_table.py}
% \rotate
 \caption{Flaring single star-planet system parameters. Obs. time is the total observing time with Kepler and TESS. \textcolor{red}{The digits in square brackets refer to the uncertainties in the corresponding last digits shown, e.g., $0.051[1]$ is the same as $0.051\pm0.001$, and $98[12]$ is the same as $98\pm12$.}}
\movetableright=-40mm

@@

The Transiting Exoplanet Survey Satellite~\citep[TESS,][]{ricker2015transiting} is an all-sky mission that began operations in 2018, and has completed two sky scans in summer 2022. It is observing at the time of writing, collecting nearly continuous photometric time series in the $600-1000\,$nm band for $\sim 27\,$d in each observing Sector. About $200\,000$ stars have been observed in 2-min cadence in the first two years of operations alone, with about $20,000$ targets per Sector. Out of these, from Sector 27 on, $1,000$ targets were observed at even higher 20-s cadence in each Sector. 
% 20000 per sector does not add up, because of stars observed in multiple sectors, the 20000 and 1000 figures can be found here: https://tess.mit.edu/observations/target-lists/
% extended TESS mission https://heasarc.gsfc.nasa.gov/docs/tess/extended.html

@@

% Mid-to-late M dwarfs follow this bimodal distribution

@@

The methods for light curve de-trending and flare finding in Kepler and TESS light curves, as well as the Anderson-Darling test, are the same as detailed in~\citet{ilin2022searching}. We briefly recap the techniques in Sections~\ref{sec:methods:flaresearch} and \ref{sec:methods:adtest}. The expected power of SPI depends on the relative velocity between the planet and the magnetic field strength in its orbit, the derivation of which we explain in 
Sections~\ref{sec:methods:relvel} and \ref{sec:methods:bfield}, respectively. We can then combine them with the stellar radius, planetary radius, and semi-major axis to estimate the power of SPI. We use the scaling laws for two different magnetic SPI models, which we introduce in Section~\ref{sec:methods:pspi}.

\subsection{Light curve de-trending and flare search}
\label{sec:methods:flaresearch}
To remove trends and rotational variability from the light curve without losing the flare signal, we use an empirically derived multistep process, implemented as the \texttt{custom\_detrending} method in \texttt{AltaiPony}, a flare science Python package for light curve analysis~\citep{ilin2021altaipony}. First, we apply a spline fit with a coarse sampling of 30h-averaged values to capture slow rotation with periods above multiple days and non-periodic trends. Then, we iteratively fit a series of sines to capture rotational signal on time scales down to about half a day, close to the fastest rotational signals measured in low mass stars. Eventually, we apply two Savitzky-Golay~\citep{savitzky1964smoothing} filters in sequence, with window sizes of 6h and 3h each. At this stage, we mask all data points above a $2.5 \sigma$ threshold (or $1.5 \sigma$ for stars like AU Mic or Proxima Cen, which are very active or have very low noise levels, or both) as flare candidates to prevent the filter from ironing out the flares. As a final step, we fit exponential functions to the edges of the light curves, if there are data points that deviate more than one standard deviation from the median value, while keeping the flare candidates masked. Transits are usually too shallow to affect either light curve de-trending or flare finding. The light curve portions around deep transits were inspected manually.

@@

We vet all flare candidates by eye, and exclude instrumental and physical false positives, such as Solar System Objects passing by the field of view~(see Fig.~\ref{fig:flares_and_fps} in the Appendix for an example). For each confirmed flare candidate, we calculate the relative amplitude $a$ and equivalent duration $ED$, defined as the flare flux $F_{\rm flare}$ over the duration of the flare, divided by the median flux $F_0$ of the star, integrated over the flare duration~\citep{gershberg1972results}:
\begin{equation}
\label{eq:ED}
ED=\displaystyle \int \mathrm dt\, \frac{F_{\rm flare}(t)}{F_0}.
\end{equation}
In other words, equivalent duration is the time during which the non-flaring star releases as much energy as the flare.

\subsection{Custom Anderson-Darling test}
\label{sec:methods:adtest}
For each star-planet system with three or more flares in its light curves, we test for deviations from a random distribution of flares with orbital phase. We use the same customized Anderson-Darling test as in \cite{ilin2022searching}, which is more sensitive to deviations at both ends of the distribution than the widely used Kolmogorov-Smirnov test~\citep{feigelson2012modern}. In short, we first take the number of flares observed in the TESS and Kepler data to calculate a base flare rate for each light curve. We then calculate how often each orbital phase has been observed, also for each light curve separately, to ensure that the different detection thresholds for TESS and Kepler do not bias the base rate. With the phase coverage and base flare rate combined, we can tell how many flares we would expect to see in any given phase bin if flares were randomly distributed. \textcolor{red}{The bin widths correspond to the distance between each consecutive detection of a flare in phase space.} We then aggregate this number of expected flares per orbital phase bin into one expected distribution (blue curves in Figs.~\ref{fig:cumdist_transiting} and \ref{fig:cumdist_rv}) by summing over all available light curves. As a last step, we compare the expected distribution to the observed one using an Anderson-Darling test. For this, we sample 10000 times from the expected distribution the same number of flares as we observe in each system. For instance, for HIP 67522, we sample 12 flares distributed according to its blue curve in Fig.~\ref{fig:cumdist_transiting}. For each of the 10000 samples, we compute the Anderson-Darling statistic $A^2$, which yields a distribution of $A^2$ values~\textcolor{red}{(see~Fig.~\ref{fig:ad_dist_hip} for the $A^2$ distribution derived for HIP~67522)}. In a last step, we compute $A^2$ for the observed distribution, and compute the $p$-value. We repeat this test with different phase offsets to account for remaining biases in sensitivity of the test at different phases.

@@

\subsection{Stellar magnetic fields}
\label{sec:methods:bfield}
We derive the average magnetic field strength $B$ from Rossby number $R$o using the empirical relation derived in \citet{reiners2022magnetism},  Table 2, in the unsaturated and saturated regimes, respectively:

@@

\begin{figure}[ht!]
    \script{paper_lx_plots.py}
    \begin{centering}
        \includegraphics[width=\linewidth]{figures/PAPER_lxlbol.png}
        \caption{
           X-ray luminosity over bolometric luminosity compared to Rossby number (\textbf{upper panel}), and average surface magnetic field strength (\textbf{lower panel}). Blue lines denote the error bars, black lines point to the IDs of the corresponding stars. Both relations follow those of stars not selected for hosting close-in planets, except for GJ 3082, which is underluminous for its Rossby number, and hence also its estimated magnetic field appears too strong for its $L_X/L_{\rm bol}$.
        }
        \label{fig:lx}
    \end{centering}
\end{figure}

@@

First, we estimate the power of SPI generated by the stretch-and-break mechanism using the scaling relations in~\cite{lanza2012starplanet}. We adopt their Eqn.~45 for non-linear and axisymmetric magnetic fields, because the non-linear formulation does not assume a constant force-free parameter, and because only axisymmetric fields can produce modulation of flaring activity in phase with the planetary orbit. We note that the axisymmetry of stellar magnetic fields varies across stars~\citep[e.g.,][]{donati2009magnetic, donati2008largescale, morin2008largescale, morin2010largescale}, and also with the stellar activity cycle of individual stars~\citep[e.g.,][]{borosaikia2016solarlike, lehmann2021identifying}. In Eqn. 45, the power of SPI, $P_{\rm spi,sb}$, is

\begin{equation}
      P_{\rm spi,sb}=\underbrace{\frac{27\pi \lambda^2(n+1)}{16 \mu_0 n (\lambda^2 + n^2)^{1/3}}}_{\hat{C}} B_{\rm p}^{2/3} R_{\rm p}^2 B_*^{4/3}  v_{\rm rel} \left(\frac{R_*}{a}\right)^{\frac{n+11}{ 3}},
\end{equation}

where $\mu_0$ is the magnetic permeability of the vacuum, and $n$ and $\lambda$ are dimensionless constants, all of which we subsume under $\hat{C}$, together with other constant factors. $B_{\rm p}$ and $B_*$ are the planetary and stellar field strengths at their poles, respectively; $v_{\rm rel}$ is the relative velocity between the stellar rotation and the planet's orbit at the semi-major axis $a$; and $R_*$ and $R_{\rm p}$ are the stellar and planetary radii, respectively. We choose $n=0.25$ ($\lambda^2 = 1.01203$), which best reproduces observations of SPI candidate systems HD~179949, $\tau$ Boo, and HD~189733 in~\citet{lanza2012starplanet} (except for $\upsilon$ And). Since the planetary fields are unknown, we adopt $B_{\rm p}=1\,$G for simplicity. We also do not know the stellar field strength at the pole for the vast majority of the stars, so we adopt an average value of 15 per cent of the surface field strength~(see Sections~\ref{sec:methods:bfield} and~\ref{sec:discussion:as}) as a proxy.

@@

\begin{equation}
      P_{\rm spi,sb0} =\hat{C} R_{\rm p}^2 B_*^2 v_{\rm rel}  \left(\frac{R_*}{a}\right)^{\frac{n+11}{ 3}}
\end{equation}

Third, we estimate the power of SPI generated by the Alfv\'en wing mechanism ~\citep{saur2013magnetic,kavanagh2022radio},  combining Eqn.~8 for the power of interaction and Eqn.~11 for the magnetopause radius of the planet from~\citet{kavanagh2022radio} to arrive at:

\begin{equation}
    P_{\rm spi,aw} = \frac{\pi^{1/2}}{2^{2/3}} B_{\rm p}^{2/3} R_{p}^2  R_*^2  B_*^{1/3}  v_{\rm rel}^2 a^{-2} \rho_*^{1/2} \sin^2 \theta
\end{equation}
